---
title: "Analyse Hurdle model"
author: "Thomas Klebel"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(brms)
library(tidybayes)
library(marginaleffects)
library(scales)
library(modelr)

# Use the Johnson color palette
clrs <- MetBrewer::met.brewer("Johnson")

# Tell bayesplot to use the Johnson palette (for things like pp_check())
bayesplot::color_scheme_set(c("grey30", clrs[2], clrs[1], clrs[3], clrs[5], clrs[4]))

# theme adapted from https://www.andrewheiss.com/blog/2021/11/10/ame-bayes-re-guide
theme_clean <- function(...) {
  theme_minimal(base_family = "Hind") +
    theme(strip.text = element_text(size = rel(1), hjust = 0),
          strip.background = element_rect(fill = "grey90", color = NA),
          plot.margin = margin(5, 5, 5, 5))
}

extrafont::loadfonts(device = "win")

# load the data
df <- read_csv(here::here("data/processed/multilevel_sample.csv"))
wdi <- WDI::WDI_data$country %>%
  as_tibble() %>%
  select(iso2c, region, income)

base <- df %>%
  left_join(wdi, by = c("country_code" = "iso2c")) %>%
  select(institution_id, University, country, region, author_position,
         P_top10, field,
         APC_in_dollar, total_weight) %>%
  mutate(APC_in_dollar = case_when(is.na(APC_in_dollar) ~ 0,
                                   TRUE ~ APC_in_dollar)) %>% 
  mutate(P_top10 = scale(log(P_top10), scale = FALSE))
```

```{r}
hm <- brm(file = "final_models/mix_final_vienna.rds", file_refit = "never")
```

# Posterior predictive check

```{r pp_check}
pp_check(hm, ndraws = 10)
```

```{r}
pred_vis <- function(df, model, country_selection, alpha = 1, ndraws = 1000) {
  get_back <- function(df) mutate(df, P_top10 = exp(6.09 + P_top10))
  
  df %>%
    filter(country == country_selection) %>%
    data_grid(P_top10, country, field) %>%
    add_predicted_draws(model, ndraws = ndraws, re_formula = NULL) %>%
    get_back() %>% 
    ggplot(aes(P_top10, .prediction)) +
    stat_interval() +
    scale_color_manual(values = colorspace::lighten(clrs[4], c(.8, .67, .42))) +
    scale_y_continuous(labels = dollar) +
    geom_point(aes(y = APC_in_dollar), alpha = alpha,
               data = filter(df, country == country_selection) %>% get_back()) +
    facet_wrap(vars(field)) +
    labs(y = "Predicted vs. actual APC", x = expression(P["top 10%"]),
         color = "Credible interval") +
    # theme_minimal(base_family = "Hind") +
    theme_clean() +
    theme(legend.position = "bottom", panel.grid.minor = element_blank())
}
```

```{r pp_austria, fig.width=10, fig.height=7}
pred_vis(base, hm, "Austria")
```

```{r pp_brazil, fig.width=10, fig.height=7}
pred_vis(base, hm, "Brazil", alpha = .4)
```
This updated model fares much better for Brazil. The predictions are still not
ideal (underestimating higher APCs of ~2000), but overall much better than the
previous model. 

```{r pp_china, fig.width=10, fig.height=7}
pred_vis(base, hm, "China", alpha = .15)
```

```{r pp_us, fig.width=10, fig.height=7}
pred_vis(base, hm, "United States", alpha = .2)
```

```{r pp_turkey, fig.width=10, fig.height=7}
pred_vis(base, hm, "Turkey", alpha = .7)
```


# Model variances and covariances
```{r}
summary(hm)
```


# Marginal effects
Here we compute marginal effects manually, by making predictions for a given
x (P_top10) and then the same x * 1.01, i.e., increasing x (on the original 
scale) by 1%, and then comparing the predictions

```{r}
scale_offset <- attributes(base$P_top10)[["scaled:center"]]
x1_identity <- 20
x2_identity <- x1_identity * 1.01
x1_log <- log(x1_identitiy) - scale_offset
x2_log <- log(x2_identity) - scale_offset


contrast <- predictions(hm, 
                        newdata = datagrid(log_x = c(2, log(exp(2)*1.01)),
                                   z = 0)) %>% 
  posteriordraws() %>% 
  mutate(x = factor(log_x, labels = c("base", "step"))) %>% 
  pivot_wider(-c(predicted, conf.low, conf.high, log_x, rowid), 
              names_from = x, values_from = draw) %>% 
  mutate(contrast = step / base)
contrast
```

This guide seemed promising: https://vincentarelbundock.github.io/marginaleffects/articles/transformation.html#back-transform-lognormal-hurdle
However, all of the marginal effects do not work, because we have a logged exposure.
The slopes are correct for the logged exposure, but not correct when we
rescale (exponentiate and un-center) the exposure. Don't know how to correctly
compute those marginal effects in this instance.

This is best shown in the below: the predicted values taper off, as P_top10
increases. But the logged P_top10 has positive AMEs that increase.

```{r}
eps <- .001
predictions_data <- predictions(
  hm,
  newdata = datagrid(country = "Brazil",
                     P_top10 = seq(-3, 3, .1),
                     field = "Biology"),
  dpar = "mu",
  transform_post = exp
) 
predictions_data %>% 
  mutate(P_top10 = exp(6 + P_top10)) %>% 
  ggplot(aes(x = P_top10, y = predicted)) +
  geom_line(size = 1)

slopes_data <- comparisons(
  hm,
  dpar = "mu",
  variables = list(P_top10 = eps),
  newdata = datagrid(country = "Brazil",
                     P_top10 = c(-2, 0, 2),
                     field = "Biology"),
  # rescale the elements of the slope
  # (exp(40.001) - exp(40)) / exp(0.001)
  transform_pre = function(hi, lo) ((exp(hi) - exp(lo)) / exp(eps)) / eps
) %>% 
  left_join(predictions_data, by = "P_top10") %>%
  # Point-slope formula: (y - y1) = m(x - x1)
  mutate(intercept = comparison * (-P_top10) + predicted)

ggplot(predictions_data, aes(x = P_top10, y = predicted)) +
  geom_line(size = 1) + 
  geom_abline(data = slopes_data, aes(slope = comparison, intercept = intercept), 
              size = 0.5, color = "red") +
  geom_point(data = slopes_data) +
  geom_label(data = slopes_data, aes(label = paste0("Slope: ", round(comparison, 1))),
             nudge_x = -1, hjust = 1) +
  theme_minimal()
```


## Revert back to directly plotting effects
### Effect of ptop10 on APC
THE SCALE IS OFF HERE: IT IS NOT A 25% CHANGE, BUT A .25% CHANGE!!
```{r mu-effect-field, fig.width=7, fig.height=8}
hm %>%
  spread_draws(b_P_top10, r_field[field,term]) %>%
  filter(term == "P_top10") %>%
  mutate(total_effect = b_P_top10 + r_field) %>%
  ggplot(aes(x = total_effect, y = reorder(field, total_effect))) +
  geom_vline(xintercept = 0, linetype = 2, alpha = .5) +
  stat_halfeye() + 
  scale_x_continuous(labels = percent) +
  labs(x = "Change in APC for 1% increase in P_top10", y = NULL) +
  theme_clean()
```

```{r mu-effect-country, fig.width=7, fig.height=15}
hm %>%
  spread_draws(b_P_top10, r_country[country,term]) %>%
  filter(term == "P_top10") %>%
  mutate(total_effect = b_P_top10 + r_country) %>%
  ggplot(aes(x = total_effect, y = reorder(country, total_effect))) +
  geom_vline(xintercept = 0, linetype = 2, alpha = .5) +
  stat_halfeye() + 
  scale_x_continuous(labels = percent) +
  labs(x = "Change in APC for 1% increase in P_top10", y = NULL) +
  theme_clean()
```

Important question: is the above misfit for lower level APCs influencing these
estimates? 


## Effect on hurdle
```{r hu-effect-field, fig.width=7, fig.height=8}
hm %>%
  spread_draws(b_hu_P_top10, r_field__hu[field,term]) %>%
  filter(term == "P_top10") %>%
  mutate(total_effect = b_hu_P_top10 + r_field__hu) %>%
  ggplot(aes(x = total_effect, y = reorder(field, total_effect))) +
  geom_vline(xintercept = 0, linetype = 2, alpha = .5) +
  stat_halfeye() + 
  scale_x_continuous(labels = percent) +
  labs(x = "Change in probability of no APC for 1 unit increase in log(P_top10)", 
       y = NULL) +
  theme_clean() +
  coord_cartesian(xlim = c(-.5, .5))
```

Interpretation: For fields like mathematics or physics, an increase in P_top10
is associated with an increase in zero APCs, whereas in fields like 
enivornmental science, psychology or biology, it is associated with a decrease.
But is the %/% interpretation correct here? This seems like a strong effect.

This is the change in probability of no APC for an increase of "1" on the logged
P_top10 scale, so roughly one standard deviation. 

```{r hu-effect-country, fig.width=7, fig.height=15}
hm %>%
  spread_draws(b_hu_P_top10, r_country__hu[country,term]) %>%
  filter(term == "P_top10") %>%
  mutate(total_effect = b_hu_P_top10 + r_country__hu) %>%
  ggplot(aes(x = total_effect, y = reorder(country, total_effect))) +
  geom_vline(xintercept = 0, linetype = 2, alpha = .5) +
  stat_halfeye() + 
  scale_x_continuous(labels = percent) +
  labs(x = "Change in probability of no APC for 1% increase in P_top10", 
       y = NULL) +
  labs(x = "Change in probability of no APC for 1 unit increase in log(P_top10)", 
       y = NULL) +
  theme_clean() +
  coord_cartesian(xlim = c(-.75, .75))
```

# Marginal effects for selected countries
US, China, Brazil, Austria, France, India, South africa
show the lines here
